### 2015-05-08 tex/prelude.tex
setup with multi files

\PassOptionsToClass{11pt,withtimes}{easychair}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{scaled=.86}{helvet}
\PassOptionsToPackage{
   a4paper,
   pdftex,
   pdftitle={Model Finding for Recursive Functions in SMT},
   pdfauthor={Andrew Reynolds, Jasmin Christian Blanchette, and Cesare Tinelli},
   pdfkeywords={},
   pdfborder={0 0 0},
   draft=false,
   bookmarksnumbered,
   bookmarks,
   bookmarksdepth=2,
   bookmarksopenlevel=2,
   bookmarksopen}{hyperref}
\documentclass{easychair}